---
title: "Assignment COURSE PROJECT 1"
author: "Tamara Rueda"
date: "19 de marzo de 2018"
output: html_document
---
1.- Code to reading in the dataset
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
ACTIVITY <- read.csv("activity.csv", header = TRUE, sep = ",")
##A data.frame without NA큦
ACTIVITY <- na.omit(ACTIVITY)
```
2.- histogram of the total number of steps taken each day
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
hist(ACTIVITY$steps, main = "Total Number of Steps", breaks = "sturges", xlab = "Steps", col = c("red", "blue", "green", "darkblue", "yellow"))
```
3.- Mean and Median number of steps taken each day
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
steps <- with(ACTIVITY, tapply(steps, date, sum))
summary(steps)
```
4.- Time series plot of the average number of steps taken
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
SubsActivity <- subset.data.frame(ACTIVITY, select = steps:date)
date <- SubsActivity$date
library(lubridate)
library(plotly)
day_activity <- day(SubsActivity$date)
SubActiv.grp.day <- group_by(SubsActivity, date) # column name to group by which is date
library(dplyr)
p<- summarize(SubActiv.grp.day, mean(steps))
p <- as.data.frame(p)
colnames(p) <- c("day", "steps")
##We are plotting the daily averaged number of steps. day number of the month is "Day"
```
```{r Daily Aver Steps Plot, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
g <-with(p, plot(steps, type = "l", main = "Daily Average Steps", col = "dark blue", lwd = 3.5))
```
5.-The 5-minute interval that, on average, contains the maximum number of steps
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
dailymeans <- with(ACTIVITY, tapply(steps, interval, mean))
dailymeans[which(dailymeans == max(dailymeans))]##to find max value of steps and which interval is
```
6.-Code to describe and show a strategy for imputing missing data
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
#read again the file now with the NA큦 
dat <- read.csv("activity.csv", head = TRUE, sep = ",", na.strings = "")
#now NA큦 are replaced with the mean value
dat$steps <- ifelse(is.na(dat$steps), mean(dat$steps, na.rm=TRUE), dat$steps)
head(dat)
```
7.- Histogram of the total number of steps taken each day after missing values are imputed
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
hist(dat$steps, breaks = 5, main = paste("Hist.steps, mean instead of NA큦"), freq = TRUE, col = c("red", "blue", "yellow", "green", "brown", "purple"))
```
8.- Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(plyr)
#mutate to modify a data.frame adding or replacing columns
#weekdays function will add a column with corresponding weekdays in spanish in my case
ACTIVITY$date <- as.Date(ACTIVITY$date)
activmod <- mutate(ACTIVITY, day = weekdays(ACTIVITY$date))
weekend <- c("sabado", "domingo")#beware spanish weekend days
weekday <- c("lunes", "martes", "miercoles", "jueves", "viernes")
activmod$day <- weekdays.Date(activmod$date)#columna day contiene dia de la semana
wkds <- subset(activmod, day %in% weekday)
weekdaymeans <- with(wkds, tapply(steps, interval, mean))
##Now weekend Days
weekends <- subset(activmod, day %in% weekend)
weekendsDayMeans <- with(wkds, tapply(steps, interval, mean))

par(mfrow = c(1,2))
plot(weekdaymeans, type = "l", col = "dark blue", main = "Aver Steps Weekdays", xlim=c(0,300), ylim=c(0,300))
plot(weekendsDayMeans, type = "l", col = "green", main = "Aver Steps Weekends", xlim=c(0,300), ylim=c(0,300))


```